TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-04-26 paper/draft.tex
note to JB

\begin{abstract}
We present a catalog of spectroscopic temperatures, metallicities,
spectral types for 10,311 M dwarfs in the \apogee\ survey using
\thecannon, which is a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs in the temperature range $2860
< \teff < 4130$\,K, and metallicty range $-0.5 < \feh < 0.5$\,dex
[WITH LABELS FROM WHERE??], we train a two-parameter model with
predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex
respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use empirically callibrated relations to estimate masses
radii from $M_K$ absolute magnitudes and \cannon\ temperatures.
This project has implications for the NASA \tess\ Mission and other
surveys that depend on having good parameters for M dwarfs.
\end{abstract}

>>>>

\begin{abstract}
We present a catalog of spectroscopic temperatures, metallicities,
spectral types for 10,311 M dwarfs in the \apogee\ survey using
\thecannon, which is a flexible, data-driven spectral-modeling and
parameter-inference framework demonstrated to estimate
stellar-parameter labels ($\teff$, $\logg$, $\feh$, and detailed
abundances) to high precision.
Using a training sample of 87 M dwarfs in the temperature range $2860
< \teff < 4130$\,K, and metallicty range $-0.5 < \feh < 0.5$\,dex
[WITH LABELS FROM WHERE??], we train a two-parameter model with
predictive accuracy (in cross-validation) to 77\,K and 0.09\,dex
respectively.
We also train a one-dimensional spectral type model using 51 M dwarfs
with known [FROM WHERE??] spectral types ranging M0 to M9, to predictive accuracy of
0.9 types.
From the overlap of \textsl{APOGEE DR14} and \textsl{Gaia DR2} we
compare our \cannon\ derived $\teff$ against color--temperature
relations and \cannon\ derived $\feh$ against isochrone models in
\gaia\ color-magnitude space.
Finally we use empirically callibrated relations to estimate masses
radii from $M_K$ absolute magnitudes and \cannon\ temperatures.
This project has implications for the NASA \tess\ Mission and other
surveys that depend on having good parameters for M dwarfs.
\end{abstract}
